### 2014-05-30 ACL2014/abstract.tex
Updated latex files

\begin{abstract}
Millions of users visit commercial search engines and  ``query'' their interests.
To provide users with high quality of services, 
search engines such as Bing, Google, and Yahoo require
intelligent analysis to realize users' implicit intents, in particular taking
advantage of very large scaled query logs.  One of the key
interesting tasks involved in learning users' implicit intents often involves 
computing pairwise similarity between all queries.
Due to the large scaled volume that is based on streaming querying, 
computing pairwise similarity is a computationally time intensive task. 
To challenge the aforementioned problem, we exploit  several
Locality Sensitive Hashing (a.k.a, LSH) methods and their novel variants. 
In particular, we develop the framework of LSHs on a distributed system (e.g. Hadoop) to take
advantage of its computing efficiency.

\end{abstract}

>>>>

\begin{abstract}
Many Web, Mobile, and NLP applications 
rely on finding nearest neighbors 
for a given query from a commercial large-scale query logs . 
Due to the large scale volume of queries, 
computing nearest neighbors from all queries using brute force linear 
search is a computationally time intensive task. 
To challenge the aforementioned problem, we exploit existing 
\lshf (a.k.a, \lsh) methods and propose 
their novel variants in distributed setting (e.g. Hadoop). 
Our experimental results show that couple of our 
new variants of \lsh get signiﬁcantly better recall than vanilla \lsh by 
using the same number of hash tables. 
\end{abstract}

### 2014-05-30 ACL2014/intro.tex
Updated latex files

 \section{Introduction}
Many real-world problems on the Web like spelling correction \cite{}, 
finding related queries \cite{Jones06WWW,Jain11SIGIR,Song12WSDM}, finding near-duplicate queries \cite{}, 
and diversifying search results \cite{Song11SIGIR} 
can benefit from query logs from a commercial search engine. 
We can take advantage of large-scale query logs 
by computing pairwise similarity between all queries. Each query point also has an associated feature vector, that is
very high dimensional and sparse $\--$ these are the set of webpages (urls) that
get clicked for this query term.

However, computing pairwise similarity between 
all queries on a commercial large-scale query logs (consisting of hundreds of millions of  queries) 
is a computationally time intensive task. 
To solve this problem, we exploit research advances in Locality Sensitive Hashing 
\cite{Indyk98STOC,Charikar02STOC,Andoni06FOCS,Andoni08CACM}  
to propose an efficient approximate similarity Hadoop framework.

\begin{itemize}
\item We present a distributed system that can approximately compute pairwise similarity 
between all queries on a commercial large-scale query logs in a time efficient manner. 
Can we claim something more specific wrt time (e.g. 0(N logN))?
\item Vanilla LSH 
\item We propose two novel variants of vanilla LSH motivated by the research on Multi-Probe LSH \cite{LvVLDB07}: Random Flipping and Query-Driven Flipping. 
\item  We show that the applicability of our system on two real-world applications like finding related queries (Increasing  direct display coverage is related to finding related queries) and ??.
\end{itemize}

@@

Spelling {grouping mis-spellings in same group}
Similar entities {LOTR,}
Finding related queries

>>>>

 \section{Introduction}
Millions of users visit commercial search engines and ``query'' 
their interests. To provide users with high quality of services, 
search engines such as Bing, Google, and Yahoo require
intelligent analysis to realize users' implicit intents, in particular taking
advantage of very large scaled query logs.  One of the key
interesting tasks involved in learning users' implicit intents often involves 
computing nearest neighbors (queries) for a user given query from all queries. 
Computing nearest neighbors is useful for 
many search-related problems on the Web and the Mobile such as 
finding related queries \cite{Jones06WWW,Jain11SIGIR,Song12WSDM}, 
finding near-duplicate queries \cite{}, spelling correction \cite{},  
, paraphrasing \cite{petrovicNAACL12,ganitkevitch13Paraphrase}, 
and diversifying search results \cite{Song11SIGIR}.

In this research,  we will take advantage of large-scale query logs 
by finding nearest neighbors (queries) for a user given query from all 
queries from a big query logs. Each query point also has an associated feature vector, that is
very high dimensional and sparse $\--$ these are the set of webpages (urls) that get clicked for this query term.

However, computing nearest neighbors (queries) for a user given query 
from a commercial large-scale query logs (consisting of hundreds of millions of  queries) is not scalable. 
The reason being conducting pairwise computations between a given query and all the queries in the dataset involves 
linear number of comparisons in the query dataset size. For finding neighbors for hundreds of millions of queries, 
it means doing petabytes of comparisons; that is computationally challenging or infeasible even in a distributed setting (such as Hadoop). 
In our experiments, our naive Hadoop implementation could not compute neighbors of 
2000 randomly sampled queries over the commercial query-logs 
consisting of hundreds of millions of queries.

The motivated behind this research is to find nearest neighbors by doing a 
\emph{small} number of comparisons (sublinear in dataset size), instead of brute force linear search. 
In addition to \emph{small} number of comparisons, we also want to 
return a set of neighbor candidates with a $100\%$ precision and a large recall. 
The method we propose meet all these criterions. We do this by exploiting 
existing research in Locality Sensitive Hashing 
\cite{Indyk98STOC,Charikar02STOC,Andoni06FOCS,Andoni08CACM} and propose their 
novel variants. In particular, we develop the framework of  \lsh algorithms on a distributed 
system (e.g. Hadoop) to take advantage of its computing efficiency.

Our work includes following contributions: 
\begin{enumerate}
\item We present a distributed system (Hadoop) that can approximately find nearest neighbors for a given query from a commercial large-scale query logs in sublinear time. % Can we claim something more specific wrt time (e.g. 0(N logN))?
\item We present vanilla \lsh algorithm based on the seminal research of Andoni and Indyk  \shortcite{Andoni08CACM}. To best of our knowledge, this is the first paper that applies this algorithm to NLP applications. 
\item We propose four novel variants of vanilla LSH motivated by the research on Multi-Probe LSH \cite{LvVLDB07}.  We show that two of our variants get signiﬁcantly better recall than the vanilla \lsh by using the same number of hash tables. The main idea behind these variants is to intelligently look up multiple buckets within a table that have a high probability of containing the nearest neighbors of a query.  
%: Random Flipping and Query-Driven Flipping. 
\item  We show that the applicability of our system on two real-world applications like finding related queries (Increasing  direct display coverage is related to finding related queries) and ??.
\end{enumerate}

@@

%Spelling {grouping mis-spellings in same group}
%Similar entities {LOTR,}
%Finding related queries

### 2014-05-30 ACL2014/system.tex
Updated latex files

This algorithm is exact as it has both $100\%$ precision and recall. 
However, the algorithm is naive as it is not scalable for hundreds of millions of queries. 
The reason being conducting pairwise computations between all the queries in the dataset involves 
quadratic number of comparisons in the query dataset size. 
For hundreds of millions of queries, it means doing petabytes of comparisons; 
that is be computationally challenging or infeasible even in a distributed setting (such as Hadoop).

@@

A family is generally interesting when  $P_1>P_2$. However, the difference between $P_1$ and $P_2$ can be very small. 
Given a family $H$ of hash functions with parameters ($R, cR, P_1,P_2$), 
\lsh algorithm is devised by amplifying the grap between the two probabilities $P_1$ and $P_2$. 
The pseudo code written in distributed framework (Hadoop) for vanilla \lsh is shown below:

@@

>>>>

This algorithm is exact as it has both $100\%$ precision and recall. 
However, the algorithm is naive as it is not scalable for hundreds of millions of queries. 
The reason being conducting pairwise computations between all the queries in the dataset involves 
quadratic number of comparisons in the query dataset size. 
For hundreds of millions of queries, it means doing petabytes of comparisons; 
that is  computationally challenging or infeasible even in a distributed setting (such as Hadoop).

@@

A family is generally interesting when  $P_1>P_2$. 
However, the difference between $P_1$ and $P_2$ can be very small. 
Given a family $H$ of hash functions with parameters ($R, cR, P_1,P_2$), 
\lsh algorithm is devised by amplifying the gap between 
the two probabilities $P_1$ and $P_2$ by concatenating several functions. 
In particular, \lsh algorithm concatenates $K$ hash functions: 
$g(q)=(h_1(q),h_2(q),\dots,h_k(q))$. The larger value of $K$ leads to larger 
gap between probabilites of collision between close queries and far queries; 
the probabilities are $P_1^{K}$ and $P_2^{K}$ respectively. 
The advantage of this amplication is that now the algorithm is more precise meaning 
the probabilty of similar queries falling in same bucket is big; 
and dis-similar queries falling in same bucket is small.

In addition to the \lsh algorithm being precise, it needs to have large recall. 
Instead of generating one set of hash tables with index $g(q)=(h_1(q),h_2(q),\dots,h_k(q))$, 
\lsh algorithm generates $L$ tables, $g_j(q)=(h_{1,j}(q),h_{2,j}(q),\dots,h_{k,j}(q)))$; $\forall 1 \leq j \leq  L$.

%Recall = $1-(1-P_1^{K})^{L}$

The pseudo code written in distributed framework (Hadoop) for vanilla \lsh is shown below:

@@

In this paper, we are interested in cosine similarity, and use the \lsh family defined by Charikar \shortcite{Charikar02STOC}. 
For two queries; $p,q \in \mathbb{R}^D$, the cosine similarity is $\left( \frac{p.q}{ \| p \| \| q \| }\right)$. 
The \lsh functions for cosine similarity is defined using sign($\alpha.p$); where $\alpha \in \mathbb{R}^D$ is a random vector. 
Negative sign is represented as zero and positive sign as one; hence indexes of buckets in hash tables are $K$-sized bit signatures.  
To generate $\alpha$, we exploit the prior work on 
generating random projections \cite{Achlioptas03,LiH06} and 
use hash functions trick to implictly map features ($\mathbb{R}^D$) to a set of \{-1,+1\}.
\footnote{Note, we respresent our queries ($p$ and $q$) using only non-zero features.}  
This hashing trick is important and useful as we do not need to explictly store the 
huge random projection matrix of size $D \times K \times L$.
